A Spatial AR System for Wide-area Axis-aligned Augmentation of Planar Scenes in Industrial Settings

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen. In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head mounted display. However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene. For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.

@@

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head mounted display. Reliance on head mounted displays has two adverse consequences: (i) a head mounted display must be worn by each individual wishing to view the augmentation, and (ii) such glasses---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems. Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the image will appear distorted). Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction, often using software bundled with the projector.

@@

The hardware setup employed in this work comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute. In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. The floorspace used for our experiments measured dimensions of X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.

@@

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X}$ to its corresponding pixel~$\mathbf{x}$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection~$\mathbf{C}$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.

In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix $\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}. To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences. Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to camera calibration concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene. A convenient consequences of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.

\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface of known scale to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.

\paragraph{Projector calibration} As in the case of camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles. We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, giving the 2D positions of our 2D-3D correspondences. We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each. Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern; given the scene plane (a point to which we shall return in the paragraph that follows), then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane. Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

>>>>

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen. In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display. However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene. For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.

@@

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display. Reliance on head-mounted displays has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems. Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the image will appear distorted). Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction, often using software bundled with the projector.

@@

The hardware setup employed in this work comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute. In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.

@@

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection~$\mathbf{C} \in \mathbb{R}^3$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.

In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}. To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences. Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to camera calibration concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene. A convenient consequences of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.

\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.

\paragraph{Projector calibration} As in the case of camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles. We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, giving the 2D positions of our 2D-3D correspondences. We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each. Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern; given the scene plane (the recovery of which we shall turn to in the paragraph that follows), then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane. Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
